---
title: "Cell communication"
output: html_document
date: "2024-02-22"
---

# cellchat 
``` {r}
devtools::install_github("sqjin/CellChat")
library(tidyverse)
library(CellChat)
library(Seurat)
library(reticulate)

#read cell chat objects back in ! 
saveRDS(cellchat, file = "cellchat.rds")
cellchat<- read_rds("cellchat.rds")
saveRDS(cellchat.first.art, file = "cellchat.first.rds")
saveRDS(cellchat.second.art, file = "cellchat.second.rds")

``` 

# load seurat obj
``` {r}
seurat.obj.cellchat<- readRDS("all.clusters.annotated.obj.rds")
DimPlot(seurat.obj.cellchat, reduction = "umap", label = TRUE, label.size = 3.5) +
  theme_minimal() +
  NoLegend()
view(seurat.obj@meta.data)
``` 


#Subset ICI arthritis! 
``` {r}
ICI_arthritis<- subset(seurat.obj, subset = type %in% c("First arthritis")) 
``` 


# Create CellChat obj! 
``` {r}
cellchat<- createCellChat(object = seurat.obj.cellchat)
```


``` {r}
CellChatDB <- CellChatDB.human # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)
``` 
``` {r}
# Load the ligand-receptor interaction database
CellChatDB.human <- CellChatDB.human # Load the human database 
CellChatDB.mouse <- CellChatDB.mouse # Load the mouse database
CellChatDB.zebrafish <- CellChatDB.zebrafish # Load the zebrafish database
``` 


# Show the ligand-receptor categories
``` {r}
showDatabaseCategory(CellChatDB.human)
``` 

``` {r}
showDatabaseCategory(CellChatDB.mouse)
``` 
``` {r}
showDatabaseCategory(CellChatDB.zebrafish)
``` 
# Set all CellChatDB for cell-cell communication analysis
``` {r}
CellChatDB.use <- CellChatDB.human
``` 

# Set a subset of CellChatDB for cell-cell communication analysis
``` {r}
CellChatDB.use <- subsetDB(CellChatDB.human, search = "Secreted Signaling")
CellChatDB.use <- subsetDB(CellChatDB.human, search = "ECM-Receptor")
CellChatDB.use <- subsetDB(CellChatDB.human, search = "Cell-Cell Contact")
``` 

# Update CellChatDB
``` {r}
interaction <- CellChatDB.human$interaction
complex <- CellChatDB.human$complex
cofactor <- CellChatDB.human$cofactor
geneInfo <- CellChatDB.human$geneInfo
``` 

``` {r}
write.csv(interaction,  file = "interaction.csv")
write.csv(complex,  file = "complex.csv")
write.csv(cofactor,  file = "cofactor.csv")
write.csv(geneInfo,  file = "geneInfo.csv")
```  

# Adding users’ curated ligand-receptor pairs
# Update
``` {r}
options(stringsAsFactors = FALSE)
interaction <- read.csv(file = 'interaction.csv', row.names = 1)
complex <- read.csv(file = 'complex.csv', row.names = 1)
cofactor <- read.csv(file = 'cofactor.csv', row.names = 1)
geneInfo <- read.csv(file = 'geneInfo.csv', row.names = 1)
CellChatDB.human.updated <- list()
CellChatDB.human.updated$interaction <- interaction
CellChatDB.human.updated$complex <- complex
CellChatDB.human.updated$cofactor <- cofactor
CellChatDB.human.updated$geneInfo <- geneInfo
CellChatDB.use <- CellChatDB.human.updated
``` 

# Inference of Cell-cell Communication Network
``` {r}
library(CellChat)
library(patchwork)
library(circlize)
options(stringsAsFactors = FALSE)
``` 

# 2.Add CellChatDB in your cellchat object
# Set the full human ligand-receptor interaction database

``` {r, fig.width= 13, fig.height= 8}
CellChatDB.use <- CellChatDB.human

# Using a subset of CellChatDB for cell-cell communication analysis
CellChatDB.use <- subsetDB(CellChatDB.human, search = "Secreted Signaling")

# Add the Secreted Signaling database in the CellChat object
cellchat@DB <- CellChatDB.use

# 3.Subset and pre-processing the expression data 
# subset the expression data to use less RAM
cellchat <- subsetData(cellchat)

# Pre-processing the expression data
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)

# Optional: project gene expression data onto protein-protein interaction (PPI)
 cellchat <- projectData(cellchat, PPI.human)

# 4. Compute the communication probability and infer cellular communication network
cellchat <- computeCommunProb(cellchat) 

# 5. Filter out the cell-cell communication if there are only few number of cells 
# in certain cell groups
cellchat <- filterCommunication(cellchat, min.cells = 10)

# 6. Infer the cell-cell communication at a signaling pathway level
cellchat <- computeCommunProbPathway(cellchat)

# 7. Calculate the aggregated cell-cell communication network
cellchat <- aggregateNet(cellchat)
cellchat@net$count
cellchat@net$weight

# 8. visualize the aggregated cell-cell communication network
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1, 2), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, 
   weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, 
   weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")


``` 





``` {r, fig.width= 13, fig.height= 8,  message=FALSE, warning=FALSE}
#  examine the signaling sent from each cell group
mat <- cellchat@net$weight
par(mfrow = c(2, 6), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = F, 
                   edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
``` 


``` {r, fig.width= 13, fig.height= 8, message=FALSE, warning=FALSE }
mat <- cellchat@net$weight
par(mfrow = c(1, 2), xpd=TRUE)
for (i in 1:nrow(mat)) {  
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))  
  mat2[i, ] <- mat[i, ]  
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, 
                   edge.weight.max = max(mat), title.name = rownames(mat)[i])}

``` 

# Cellchat pathway! 
``` {r, message=FALSE, warning=FALSE}
cellchat@netP[["pathways"]]
``` 
``` {r}
extractEnrichedLR(cellchat, signaling = c(cellchat@netP[["pathways"]]),
                  geneLR.return = TRUE)


``` 


``` {r, fig.width= 13, fig.height= 10}
# visualize the contribution of each LR pairs to the communication network
netAnalysis_contribution(cellchat, 
                         signaling = c(cellchat@netP[["pathways"]]), 
                         title = "Contribution of each LR pairs")
``` 

``` {r}
netAnalysis_contribution(cellchat, 
                         signaling = c(cellchat@netP[["pathways"]][1:5]), 
                         title = "MIF,GALECTIN,CXCL,IL2,COMPLEMENT")
``` 
``` {r}
extractEnrichedLR(cellchat, signaling = "TNF", geneLR.return = FALSE)
netAnalysis_contribution(cellchat, signaling = "TNF")
``` 
``` {r}
extractEnrichedLR(cellchat, signaling = "CCL", geneLR.return = FALSE)

``` 
``` {r}
netAnalysis_contribution(cellchat, signaling = "CCL")

```

# Circle plot
``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
netVisual_aggregate(cellchat, signaling = "TNF", layout = "circle")
```

``` {r, message=FALSE, warning=FALSE, fig.width= 10, fig.height= 6}
netVisual_individual(cellchat, signaling = "TNF", layout = "circle")
```

``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
netVisual_individual(cellchat, signaling = "TNF", 
                     pairLR.use = "TNF_TNFRSF1A",
                     layout = "circle")
``` 
# Chord diagram

``` {r, fig.width= 13, fig.height= 9}
par(mfrow = c(1, 1), xpd=TRUE)
par(cex = 0.5)
netVisual_aggregate(cellchat, signaling = "TNF", layout = "chord")
```

``` {r, fig.width= 9, fig.height= 8, message=FALSE, warning=FALSE}
netVisual_chord_cell (cellchat, signaling = "TNF")

```

``` {r, fig.width= 13, fig.height= 9, message=FALSE, warning=FALSE}
netVisual_chord_gene (cellchat, signaling = "CXCL")
``` 

# Chord diagram: group cell clusters into fibroblast, DC and TC cells # error ! 

``` {r}
group.cellType <- c(rep("mDc", 4), rep("pDC", 4), rep("MAIT", 4))
names(group.cellType) <- levels(cellchat@idents)
par(mfrow = c(1, 1), xpd=TRUE)
par(cex = 0.5)
netVisual_chord_cell(cellchat, signaling = "TNF", 
    group = group.cellType, 
    title.name = paste0("TNF_", "signaling network"))
``` 


# Chord diagram: define source and target cell types
``` {r, fig.width= 10, fig.height= 5, message=FALSE, warning=FALSE}
netVisual_chord_gene(cellchat, sources.use = 4, targets.use = c(0:21), 
                     lab.cex = 0.5,legend.pos.y = 30)
``` 
``` {r, fig.width= 10, fig.height= 7, message=FALSE, warning=FALSE}
netVisual_chord_gene(cellchat, sources.use = c(1,2,3,4), targets.use = 8,
                     lab.cex = 0.5, legend.pos.x = 15)
``` 

# Chord diagram: show LR pairs associated with certain signaling pathways
``` {r,  fig.width= 8, fig.height= 6}
netVisual_chord_gene(cellchat, sources.use = c(1,2,3,4), targets.use = 8,
                     signaling = c("CCL","CXCL"),legend.pos.x = 8)
``` 
# Hierarchy plot  # error===> no significant communication of CXCL
``` {r, message=FALSE, warning=FALSE, fig.width= 7, fig.height= 6}
vertex.receiver = seq(1,4) # define the left portion of cell groups  
netVisual_aggregate(cellchat, signaling = "TNF", 
           vertex.receiver = vertex.receiver, layout = "hierarchy")
```

``` {r}
netVisual_individual(cellchat, signaling = "CXCL", 
                     pairLR.use = "CXCL12_CXCR4", 
                     vertex.receiver = vertex.receiver, 
                     layout = "hierarchy")
```

# Heatmap
``` {r, fig.width= 8, fig.height= 5, message=FALSE, warning=FALSE}
netVisual_heatmap(cellchat, signaling = c("CXCL"), color.heatmap = "Reds")
``` 
# Violin plot 
``` {r, fig.width= 8, fig.height= 5, message=FALSE, warning=FALSE}
plotGeneExpression(cellchat, signaling = "TNF")
``` 

# bubble plot: show all LR pairs from source to target cell groups
``` {r, message=FALSE, fig.width= 8, fig.height= 5, message=FALSE, warning=FALSE}
ICI_arthritis<- netVisual_bubble(cellchat, sources.use = 5, targets.use = c(0:21),
                 remove.isolate = FALSE, 
                 font.size = 9)
ICI_arthritis
``` 

``` {r}
ggsave("ICI_arthritis.pdf", plot = ICI_arthritis, height=5, width = 6, units = "in", dpi = 300)

``` 

``` {r, message=FALSE, fig.width= 8, fig.height= 5, message=FALSE, warning=FALSE}
netVisual_bubble(cellchat, sources.use = 5, targets.use = c(0:21), 
                 signaling = c("CCL","CXCL"), remove.isolate = FALSE) + 
  theme_get() +
  theme(axis.text.x = element_text(angle = 70, vjust = 1, hjust = 1))
``` 



``` {r, fig.width= 8, fig.height= 5, message=FALSE, warning=FALSE}
# bubble plot: show LR pairs associated with certain signaling pathways
netVisual_bubble(cellchat, sources.use = 5, targets.use = c(5:12), 
                 signaling = c("CCL","CXCL"), remove.isolate = FALSE)




``` 

``` {r, fig.width= 10, fig.height= 11 , message=FALSE, warning=FALSE}
# Run the function to generate the plot and assign it to an object
dev.new()
pdf(file ="cellchat.pdf", width = 20, height =16)
netVisual_chord_gene(cellchat, 
                          sources.use = 5, 
                          targets.use = c(0:21), 
                          lab.cex = 0.7, 
                          legend.pos.y = 0,
                          small.gap = 2,
                     show.legend = FALSE)

dev.set(dev.next())
ggsave("cellchat.tiff", plot = p, height=3, width = 5, units = "in", dpi = 300)
``` 

``` {r}
ggsave("net.pdf", plot = net_plot, height=5, width = 8, units = "in", dpi = 300)
``` 

# Systematic Analysis of Cell-cell Communication Networks 
``` {r, message=FALSE}
library(CellChat)
library(NMF)
library(ggalluvial)

# 1. Compute the network centrality scores
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")

# Scatter plot to visualize aggregated communication networks for each cell type
netAnalysis_signalingRole_scatter(cellchat) # all signaling pathways
``` 

``` {r, message=FALSE}
# Scatter plot to Visualize selected communication networks
netAnalysis_signalingRole_scatter(cellchat, signaling = "TNF")
```

``` {r, message=FALSE}
netAnalysis_signalingRole_scatter(cellchat, signaling = c("CXCL", "CCL"))

``` 
``` {r, message=FALSE, warning=FALSE}
# Scatter plot to Visualize selected communication networks
netAnalysis_signalingRole_scatter(cellchat, signaling = "TNF")
```

``` {r, message=FALSE, warning=FALSE}
netAnalysis_signalingRole_scatter(cellchat, signaling = c("CXCL", "CCL"))
``` 

``` {r, message=FALSE, warning=FALSE}
# Heatmap to visualize dominant cell types for each signaling pathway
netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", height = 7)
```

``` {r}
netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", height = 7)
``` 

``` {r}
# Visualize selected outgoing/incoming signals and contributing cell types
netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing",
                                      signaling = c("CXCL", "CCL"))
```

``` {r}
netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming",
                                      signaling = c("CXCL", "CCL"))

``` 

``` {r}
# Heatmap to visualize major signaling roles of different cell groups
netAnalysis_signalingRole_network(cellchat, signaling = "TNF", width = 10, 
                                  height = 5, font.size = 10)


```



``` {r,  message=FALSE, warning=FALSE}
# 2. Identify global communication patterns to explore how multiple cell types 
# and signaling pathways coordinate

# Identify and visualize outgoing communication pattern of secreting cells
selectK(cellchat, pattern = "outgoing") # infer the number of patterns, NMF
```

``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
nPatterns = 3 # a suitable number of patterns is the one begin to drop suddenly.
cellchat <- identifyCommunicationPatterns(cellchat, pattern = "outgoing",
                  k = nPatterns, width = 5, height = 9)
```

``` {r, fig.width= 6, fig.height= 5, message=FALSE, warning=FALSE}
netAnalysis_river(cellchat, pattern = "outgoing") # river plot
```

``` {r, fig.width= 8, fig.height= 5, message=FALSE, warning=FALSE }
netAnalysis_dot(cellchat, pattern = "outgoing") # dot plot
``` 

``` {r, fig.width= 6, fig.height= 5, message=FALSE, warning=FALSE}
## Identify and visualize incoming communication pattern of target cells
selectK(cellchat, pattern = "incoming")

```

``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
nPatterns = 3
cellchat <- identifyCommunicationPatterns(cellchat,pattern = "incoming", 
                  k = nPatterns, width = 5, height = 9)
``` 

``` {r, fig.width= 6, fig.height= 5, message=FALSE, warning=FALSE}
netAnalysis_river(cellchat, pattern = "incoming") # river plot
```

``` {r, fig.width= 8, fig.height= 5, message=FALSE, warning=FALSE}
netAnalysis_dot(cellchat, pattern = "incoming") # dot plot
``` 

# 3. Groups signaling pathways based on their functional/structural similarities
# Identify signaling groups based on functional similarity

``` {r}
cellchat <- computeNetSimilarity(cellchat, type = "functional")
saveRDS(cellchat, file = "cellchat.rds")
cellchat<- readRDS("cellchat.rds")
``` 



``` {r}
library(reticulate)
library(CellChat)
use_python("C:/Users/skeam/AppData/Local/Programs/Python/Python312", required = TRUE)
reticulate::py_install(packages = "numpy")
reticulate::py_install(packages = "umap-learn")

reticulate::py_install(packages = c("numpy", "umap-learn"))

reticulate::py_install(packages = "umap-learn", conda = "auto")
reticulate::install_python(version = '3.12.2') # this is the python version I installed in my computer! 
reticulate::py_module_available("umap")
reticulate::py_install(packages = "umap-learn")
cellchat <- netEmbedding(cellchat, type = "functional")
``` 

``` {r}
cellchat <- netClustering (cellchat, type = "functional", do.parallel = FALSE)
``` 

``` {r,  message=FALSE, warning=FALSE}
# Visualization in 2D-space
netVisual_embedding(cellchat, type = "functional", label.size = 3.5)
```

``` {r}
netVisual_embeddingZoomIn(cellchat, type = "functional", nCol = 2)
``` 

# Identify signaling groups based on structure similarity
# multimeric ligand-receptor complexes, soluble agonists and antagonists, 
# stimulatory and inhibitory co-ligands and co-receptors


``` {r, message=FALSE, warning=FALSE}
cellchat <- computeNetSimilarity(cellchat, type = "structural")
cellchat <- netEmbedding(cellchat, type = "structural")
cellchat <- netClustering(cellchat, type = "structural",do.parallel = FALSE)

# Visualization in 2D-space
netVisual_embedding(cellchat, type = "structural", label.size = 3.5)
```


# Visualization in 2D-space
``` {r}
netVisual_embeddingZoomIn(cellchat, type = "structural", nCol = 2)
``` 

# Cell/communication network across conditions

``` {r}
library(CellChat)
library(patchwork)
library(NMF)
library(ggalluvial)
options(stringsAsFactors = FALSE)

# Filter First arthritis from seurat obj
seurat.obj.first.art<- subset(seurat.obj.cellchat, subset = type == "First arthritis")
View(seurat.obj.first.art@meta.data)

# 1. Create the NL CellChat object 
cellchat.first.art <- createCellChat(object = seurat.obj.first.art, group.by = "ident", assay = "SCT")

# 2. Add the Secreted Signaling database in the CellChat object
CellChatDB.use <- subsetDB(CellChatDB.human, search = "Secreted Signaling")
cellchat.first.art@DB <- CellChatDB.use

# 3.Subset and pre-processing the expression data 
cellchat.first.art <- subsetData(cellchat.first.art)
cellchat.first.art <- identifyOverExpressedGenes(cellchat.first.art)
cellchat.first.art <- identifyOverExpressedInteractions(cellchat.first.art)

# 4. project gene expression data onto protein-protein interaction (PPI)
cellchat.first.art <- projectData(cellchat.first.art, PPI.human) # PPI.mouse for mouse samples

# 5. Compute the communication probability and infer cellular communication network
cellchat.first.art <- computeCommunProb(cellchat.first.art, raw.use = FALSE) 

# 6. Filter out the cell-cell communication min.cells = 10
cellchat.first.art <- filterCommunication(cellchat.first.art, min.cells = 10)

# 7. Infer the cell-cell communication at a signaling pathway level
cellchat.first.art <- computeCommunProbPathway(cellchat.first.art)

# 8. Calculate the aggregated cell-cell communication network
cellchat.first.art <- aggregateNet(cellchat.first.art)

# 9. Compute the network centrality scores
cellchat.first.art <- netAnalysis_computeCentrality(cellchat.first.art, slot.name = "netP")
```
#cellchate !! 
``` {r}
mat <- cellchat.first.art@net$weight
par(mfrow = c(2, 6), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = F, 
                   edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
``` 


``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
# 10. Identify and visualize outgoing communication pattern of secreting cells
selectK(cellchat.first.art, pattern = "outgoing") 
nPatterns = 2 
cellchat.first.art <- identifyCommunicationPatterns(cellchat.first.art, pattern = "outgoing",
                                      k = nPatterns, width = 5, height = 9)
``` 

``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
# 11. Identify and visualize incoming communication pattern of target cells
selectK(cellchat.first.art, pattern = "incoming")
nPatterns = 4
cellchat.first.art <- identifyCommunicationPatterns(cellchat.first.art,pattern = "incoming", 
                                      k = nPatterns, width = 5, height = 9)
```

```{r}
# 12. Identify signaling groups based on functional similarity
cellchat.first.art <- computeNetSimilarity(cellchat.first.art, type = "functional")
cellchat.first.art <- netEmbedding(cellchat.first.art, type = "functional")
cellchat.first.art <- netClustering (cellchat.first.art, type = "functional", do.parallel = FALSE)

```

``` {r}
# 13. Identify signaling groups based on structure similarity
cellchat.first.art <- computeNetSimilarity(cellchat.first.art, type = "structural")
cellchat.first.art <- netEmbedding(cellchat.first.art, type = "structural")
cellchat.first.art <- netClustering(cellchat.first.art, type = "structural", do.parallel = FALSE)
```

# ============================================================================================================================================
                                                 #resources from authors
                                                 
# bubble plot: show all LR pairs from source to target cell groups
``` {r, message=FALSE, fig.width= 8, fig.height= 10, message=FALSE, warning=FALSE}

library(CellChat)
library(Seurat)
library(tidyverse)
cellchat.first.art<- read_rds("cellchat.first.rds")  #####

first_IA<- netVisual_bubble(cellchat.first.art, 
                       sources.use = 3, 
                       targets.use = c(0,2, 6, 7, 11, 16, 17, 18, 21), 
                       remove.isolate = FALSE, 
                      signaling = c("CXCL", "CCL", "MIF"))
first_IA
``` 


``` {r}
ggsave("first_arthritis.pdf", plot = first_IA, height=5, width = 4, units = "in", dpi = 300)

``` 



``` {r, fig.width= 10, fig.height= 11 , message=FALSE, warning=FALSE}
# Run the function to generate the plot and assign it to an object
first_IA<- netVisual_chord_gene(cellchat.first.art, 
                          sources.use = 3, 
                          targets.use = c(0,2, 6, 7, 11, 16, 17, 18, 21), 
                          lab.cex = 0.7,
                          legend.pos.y = 5,
                     small.gap = 2, 
                     signaling = c("CXCL", "CCL", "MIF"), 
                     show.legend = FALSE)
first_IA
``` 




#filter receptors/ligand of interest !! 

``` {r, fig.width= 10, fig.height= 11 , message=FALSE, warning=FALSE}
# Run the function to generate the plot and assign it to an object
first_IA<- netVisual_chord_gene(cellchat.first.art, 
                          sources.use = 3, 
                          targets.use = c(0,2, 6, 7, 11, 16, 17, 18, 21), 
                          lab.cex = 0.6,
                          legend.pos.y = 5,
                     small.gap = 2, 
                     pairLR.use = data.frame('interaction_name'= c("MIF_CD74_CD44", "CCL5_CCR1", "CCL5_CCR5",     "CCL5_CCR3", "CXCL9_CXCR3", "CXCL10_CXCR3", "CXCL11_CXCR3")),
                     show.legend = FALSE)
first_IA
``` 



# cell communication of CXCL13-CD4 T
``` {r}
first_IA_CXCL13<- netVisual_bubble(cellchat.first.art, 
                       sources.use = 10, 
                       targets.use = c(0,2, 6, 7, 11, 16, 17, 18, 21, 20, 3, 13), 
                       remove.isolate = FALSE, 
                       signaling =c("CXCL", "CCL", "MIF"))
first_IA_CXCL13
``` 

``` {r}
first_IA_CXCL13<- netVisual_chord_gene(cellchat.first.art, 
                          sources.use = 10, 
                          targets.use = c(0,2, 6, 7, 11, 17, 18, 21, 20), 
                          lab.cex = 0.6,
                          legend.pos.y = 5,
                     small.gap = 2, 
                     pairLR.use = data.frame('interaction_name'= c("MIF_CD74_CD44", "CXCL13_CXCR3", 
                                                                   "CXCL13_CXCR5")), 
                     show.legend = FALSE, 
                     )
first_IA_CXCL13
``` 


``` {r}
ggsave("first_arthritis_cxcl13.pdf", plot = first_IA_CXCL13, height=6, width = 7, units = "in", dpi = 300)
``` 

#receptor/ligand of interest !!

``` {r}
first_IA_CXCL13<- netVisual_chord_gene(cellchat.first.art, 
                          sources.use = 10, 
                          targets.use = c(0,2, 6, 7, 11, 17, 18, 21, 20), 
                          lab.cex = 0.6,
                          legend.pos.y = 5,
                     small.gap = 2, 
                     signaling = c("CXCL", "CCL", "MIF"), 
                     show.legend = TRUE, 
                     )
first_IA_CXCL13
``` 














## CXCL13 CD4 and effector CD8 T cells !! 
``` {r}
first_IA_CXCL13<- netVisual_chord_gene(cellchat.first.art, 
                          sources.use = 10, 
                          targets.use = c(0, 3), 
                          lab.cex = 0.5,
                          legend.pos.y = 5,
                     small.gap = 2, 
                     signaling = c("CXCL", "CCL", "MIF", "TNF", "IL2", "IL1"), 
                     show.legend = TRUE)
first_IA_CXCL13
``` 



# Select second arthritis
``` {r}
# Filter First arthritis from seurat obj
seurat.obj.second.art<- subset(seurat.obj.cellchat, subset = type == "Second arthritis")
View(seurat.obj.second.art@meta.data)

# 1. Create the NL CellChat object 
cellchat.second.art <- createCellChat(object = seurat.obj.second.art, group.by = "ident", assay = "SCT")

# 2. Add the Secreted Signaling database in the CellChat object
CellChatDB.use <- subsetDB(CellChatDB.human, search = "Secreted Signaling")
cellchat.second.art@DB <- CellChatDB.use

# 3.Subset and pre-processing the expression data 
cellchat.second.art <- subsetData(cellchat.second.art)
cellchat.second.art <- identifyOverExpressedGenes(cellchat.second.art)
cellchat.second.art <- identifyOverExpressedInteractions(cellchat.second.art)

# 4. project gene expression data onto protein-protein interaction (PPI)
cellchat.second.art <- projectData(cellchat.second.art, PPI.human) # PPI.mouse for mouse samples

# 5. Compute the communication probability and infer cellular communication network
cellchat.second.art <- computeCommunProb(cellchat.second.art, raw.use = FALSE) 

# 6. Filter out the cell-cell communication min.cells = 10
cellchat.second.art <- filterCommunication(cellchat.second.art, min.cells = 10)

# 7. Infer the cell-cell communication at a signaling pathway level
cellchat.second.art <- computeCommunProbPathway(cellchat.second.art)

# 8. Calculate the aggregated cell-cell communication network
cellchat.second.art <- aggregateNet(cellchat.second.art)

# 9. Compute the network centrality scores
cellchat.second.art <- netAnalysis_computeCentrality(cellchat.second.art, slot.name = "netP")

```

``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
# 10. Identify and visualize outgoing communication pattern of secreting cells
selectK(cellchat.second.art, pattern = "outgoing") 
nPatterns = 2 
cellchat.second.art <- identifyCommunicationPatterns(cellchat.second.art, pattern = "outgoing",
                                      k = nPatterns, width = 5, height = 9)
``` 


``` {r,fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
# 11. Identify and visualize incoming communication pattern of target cells
selectK(cellchat.second.art, pattern = "incoming")
nPatterns = 4
cellchat.second.art <- identifyCommunicationPatterns(cellchat.second.art,pattern = "incoming", 
                                    k = nPatterns, width = 5, height = 9)
```


``` {r}
# 12. Identify signaling groups based on functional similarity
cellchat.second.art <- computeNetSimilarity(cellchat.second.art, type = "functional")
cellchat.second.art <- netEmbedding(cellchat.second.art, type = "functional")
cellchat.second.art <- netClustering (cellchat.second.art, type = "functional", do.parallel = FALSE)
```  

``` {r}
# 13. Identify signaling groups based on structure similarity
cellchat.second.art <- computeNetSimilarity(cellchat.second.art, type = "structural")
cellchat.second.art <- netEmbedding(cellchat.second.art, type = "structural")
cellchat.second.art <- netClustering(cellchat.second.art, type = "structural", do.parallel = FALSE)
```


# bubble plot: show all LR pairs from source to target cell groups
``` {r, message=FALSE, fig.width= 8, fig.height= 5, message=FALSE, warning=FALSE}
# bubble plot: show all LR pairs from source to target cell groups
library(tidyverse)
library(CellChat)
cellchat.second.art<- readRDS("cellchat.second.rds")

second_arthritis<- netVisual_bubble(cellchat.second.art, 
                       sources.use = 3, 
                       targets.use = c(0,2, 6, 7, 11, 16, 17, 18, 21), 
                       remove.isolate = FALSE, 
                      signaling = c("CXCL", "CCL", "MIF", "TNF"))
second_arthritis

``` 

``` {r}

ggsave("second_arthritis.pdf", plot =, height=5, width = 7, units = "in", dpi = 300)

``` 

```{r}
ggsave("seconar.pdf", plot = gg, height= 5, width = 7, units = "in", dpi= 300)

```





``` {r, fig.width= 10, fig.height= 11 , message=FALSE, warning=FALSE}
# Run the function to generate the plot and assign it to an object
second_IA<- netVisual_chord_gene(cellchat.second.art, 
                          sources.use = 3, 
                          targets.use = c(0,2, 6, 7, 11, 16, 17, 18, 21), 
                          lab.cex = 0.6,
                          legend.pos.y = 5,
                     small.gap = 2, 
                     pairLR.use = data.frame('interaction_name'= c("MIF_CD74_CD44", "CCL5_CCR1", "CCL5_CCR5",     "CCL5_CCR3", "CXCL9_CXCR3", "CXCL10_CXCR3", "CXCL11_CXCR3")),
                     show.legend = FALSE)
second_IA


``` 


# Analysis of CXCL13 CD4 T for second arthritis ! 
``` {r}

second_IA_CXCL13<- netVisual_bubble(cellchat.second.art, 
                       sources.use = 10, 
                       targets.use = c(0,2, 6, 7, 11, 16, 17, 18, 21, 20, 3, 13), 
                       remove.isolate = FALSE, 
                       signaling =c("CXCL", "CCL", "MIF"))
second_IA_CXCL13
``` 


``` {r}
second_IA_CXCL13<- netVisual_chord_gene(cellchat.second.art, 
                          sources.use = 10, 
                          targets.use = c(0,2, 6, 7, 11, 17, 18, 21, 20), 
                          lab.cex = 0.6,
                          legend.pos.y = 5,
                     small.gap = 2, 
                     pairLR.use = data.frame('interaction_name'= c("MIF_CD74_CD44", "CXCL13_CXCR3", 
                                                                   "CXCL13_CXCR5")), 
                     show.legend = FALSE, 
                     )
second_IA_CXCL13
``` 



``` {r}
ggsave("second_arthritis_cxcl13.pdf", plot = second_IA_CXCL13, height=6, width = 7, units = "in", dpi = 300)
``` 

#CXCL13 vs effector CD8 !!
``` {r}
second_IA_CXCL13<- netVisual_chord_gene(cellchat.second.art, 
                          sources.use = 10, 
                          targets.use = c(0, 3), 
                          lab.cex = 0.5,
                          legend.pos.y = 5,
                     small.gap = 2, 
                     signaling = c("CXCL", "CCL", "MIF", "TNF", "IL2", "IL1"), 
                     show.legend = TRUE, 
                     )
second_IA_CXCL13
```




# Merge object lists! 
``` {r}
object.list<- list(First_arthritis= cellchat.first.art, Second_arthritis= cellchat.second.art)
names(object.list)

#merg object list 
cellchat.compare<- mergeCellChat(object.list, add.names = names(object.list))
View(cellchat.compare@meta)
``` 

``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
rank= rankNet(cellchat.compare, mode = "comparison", stacked = T, do.stat = TRUE, 
               signaling = c("IL16", "GAS", "OX40", "MIF", "IL2", "IL4", 
                             "TGBb", "IL1", "CD40", "FASLG", "XCR", "CCL")) + 
  theme(axis.text = element_text(size= 10, face= "bold"))# 42 & 47 
rank
``` 


``` {r}
ggsave("informationflow.pdf", plot = rank, height=3, width = 6, units = "in", dpi = 300)
``` 




``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
rankNet(cellchat.compare, mode = "comparison", stacked = F, do.stat = TRUE)
``` 
``` {r}
# 2. Compare the total number of interactions and interaction strength
compareInteractions(cellchat.compare, show.legend = F, group = c(1,2), measure = "count")
```
``` {r}
strength<- compareInteractions(cellchat.compare, show.legend = F, group = c(1,2), measure = "weight")
strength<- strength + 
  theme(axis.text = element_text(size= 12, face = "bold")) 

#save 
ggsave("intraction.strength.pdf", plot = strength, height=3.5, width = 4, units = "in", dpi = 300)

``` 






``` {r, fig.width= 10, fig.height= 4.5, message=FALSE, warning=FALSE}
###### scatter plot 
# 1.Compare outgoing/incoming interaction strength for all the cell types
count.sum <- sapply(object.list, function(x) {rowSums(x@net$count) + 
    colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(count.sum), max(count.sum)) # control the dot size 
gg <- list()
for (i in 1:length(object.list)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(object.list[[i]], 
  title = names(object.list)[i], weight.MinMax = weight.MinMax)
}

patchwork::wrap_plots(plots = gg)

``` 


``` {r, fig.width= 10, fig.height= 4.5, message=FALSE, warning=FALSE}
# 2. identify signalling changes associated with one cell group 
netAnalysis_signalingChanges_scatter(cellchat.compare, idents.use = "CD4 Central Memory")
```

``` {r, fig.width= 10, fig.height= 4.5, message=FALSE, warning=FALSE}
netAnalysis_signalingChanges_scatter(cellchat.compare, idents.use = "CLEC9A+ DCs", 
                                     signaling.exclude = "MIF")
``` 

``` {r, fig.width= 14, fig.height= 6, message=FALSE, warning=FALSE}
###### Circle plots
# 1. show the number of interactions between any two cell populations 
# compute the maximum number of cells and the maximum number of interactions 
weight.max <- getMaxWeight(object.list, attribute = c("idents","count"))

par(mfrow = c(1,2), xpd=TRUE)
for (i in 1:length(object.list)) {
  netVisual_circle(object.list[[i]]@net$count, weight.scale = T, label.edge= F,
  edge.weight.max = weight.max[2], edge.width.max = 12, arrow.size = 0.1,
  title.name = paste0("Number of interactions - ", names(object.list)[i]))
}

``` 

``` {r, fig.width= 15, fig.height= 6, message=FALSE, warning=FALSE}
# 2. selected pathway
weight.max <- getMaxWeight(object.list, slot.name = c("netP"), 
                           attribute =c("CXCL"))

par(mfrow = c(1,2), xpd=TRUE)
for (i in 1:length(object.list)) {
  netVisual_aggregate(object.list[[i]], signaling = c("CXCL"), layout = "circle",
  edge.weight.max = weight.max[1], edge.width.max = 10, arrow.size = 0.05, 
  signaling.name = paste("CXCL", names(object.list)[i]))
}

``` 
``` {r, fig.width= 15, fig.height= 6, message=FALSE, warning=FALSE}
# 3. Show differential number of interactions or interaction strength among 
# different cell populations, red(increased signaling)/blue(decreased signaling)
par(mfrow = c(1,2), xpd=TRUE)
netVisual_diffInteraction(cellchat.compare, comparison = c(1, 2), measure = "count", 
                          weight.scale = T, arrow.size = 0.1)
```

``` {r, fig.width= 15, fig.height= 6, message=FALSE, warning=FALSE}
netVisual_diffInteraction(cellchat.compare, comparison = c(1, 2), measure = "weight", 
                          weight.scale = T, arrow.size = 0.1)
``` 
# We do not need this pattern of analysis ! 
``` {r}
# 4. simplify the complicated network to the cell type level
group.cellType <- c(rep("FIB", 4), rep("DC", 4), rep("TC", 4))
group.cellType <- factor(group.cellType, levels = c("FIB", "DC", "TC"))
object.list <- lapply(object.list, function(x) {
                              mergeInteractions(x, group.cellType)})
cellchat <- mergeCellChat(object.list, add.names = names(object.list))

weight.max <- getMaxWeight(object.list, slot.name = c("idents", "net", "net"), 
                           attribute = c("idents", "count", "count.merged"))
``` 

# we do not need this one! 

``` {r}
# show the number of interactions or interaction strength.
par(mfrow = c(1,2), xpd=TRUE)
for (i in 1:length(object.list)) {
     netVisual_circle(object.list[[i]]@net$count.merged, weight.scale = T, 
     label.edge= T, edge.weight.max = weight.max[3], edge.width.max = 12, 
     arrow.size = 0.1,
     title.name = paste0("Number of interactions - ", names(object.list)[i]))
}

par(mfrow = c(1,2), xpd=TRUE)
netVisual_diffInteraction(cellchat.compare, weight.scale = T, comparison = c(1, 2),
            arrow.size = 0.1, measure = "count.merged", label.edge = T)
netVisual_diffInteraction(cellchat, weight.scale = T, comparison = c(1, 2),
            arrow.size = 0.1, measure = "weight.merged", label.edge = T)

## no need this analysis 
``` 


###### Heatmap
# 1. Compare outgoing/incoming signaling associated with each cell population
# combining all the identified signaling pathways from different datasets 

``` {r, fig.width= 10, fig.height= 8, message=FALSE, warning=FALSE}
library(draw)
library(raster)
library(ComplexHeatmap)
install.packages("curve")
install.packages("raster")
install.packages("draw")
all_pathways <- union(object.list[[1]]@netP$pathways, 
                       object.list[[2]]@netP$pathways)

ht1 = netAnalysis_signalingRole_heatmap(object.list[[1]], pattern = "all", 
      signaling = all_pathways, title = names(object.list)[1],  
      width = 5, height = 11, color.heatmap = "OrRd")
ht2 = netAnalysis_signalingRole_heatmap(object.list[[2]], pattern = "all", 
      signaling = all_pathways, title = names(object.list)[2], 
      width = 5, height = 11, color.heatmap = "OrRd")
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))  #called from complex heatmap! 
``` 

``` {r, fig.width= 10, fig.height= 8, message=FALSE, warning=FALSE}
ht3 = netAnalysis_signalingRole_heatmap(object.list[[1]], pattern ="outgoing", 
      signaling = all_pathways, title = names(object.list)[1], 
      width = 5, height = 11)
ht4 = netAnalysis_signalingRole_heatmap(object.list[[2]], pattern ="outgoing", 
      signaling = all_pathways, title = names(object.list)[2], 
      width = 5, height = 11)
draw(ht3 + ht4, ht_gap = unit(0.5, "cm"))
``` 

``` {r, fig.width= 10, fig.height= 8, message=FALSE, warning=FALSE}
ht5 = netAnalysis_signalingRole_heatmap(object.list[[1]], pattern = "incoming", 
      signaling = all_pathways, title = names(object.list)[1], 
      width = 5, height = 11, color.heatmap = "GnBu")
ht6 = netAnalysis_signalingRole_heatmap(object.list[[2]], pattern ="incoming",
      signaling = all_pathways, title = names(object.list)[2], 
      width = 5, height = 11, color.heatmap = "GnBu")
draw(ht5 + ht6, ht_gap = unit(0.5, "cm"))



``` 




``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
# 2. selected pathways
par(mfrow = c(1,2), xpd=TRUE)
ht <- list()
for (i in 1:length(object.list)) {
  ht[[i]] <- netVisual_heatmap(object.list[[i]], signaling = c("CXCL"), 
  title.name = paste(c("CXCL"), "signaling ",names(object.list)[i]),
  color.heatmap = "Reds")
}
complex<- ComplexHeatmap::draw(ht[[1]] + ht[[2]], ht_gap = unit(0.5, "cm")) 
  # Add row and column annotations with borders

# Draw the complex heatmap
ComplexHeatmap::draw(complex)

```
``` {r, fig.width= 8, fig.height= 6, message=FALSE, warning=FALSE}
# 3. show differential interaction number & interaction strength using heatmap
gg1 <- netVisual_heatmap(cellchat.compare, comparison = c(1, 2), measure = "count")
gg2 <- netVisual_heatmap(cellchat.compare, comparison = c(1, 2), measure = "weight")
gg1 + gg2


``` 






``` {r, fig.width= 12, fig.height= 15, message=FALSE, warning=FALSE}
###### Bubble plots
# 1. compare communication probabilities mediated by ligand-receptor pairs from 
# selected sources and targets cell groups 
netVisual_bubble(cellchat.compare, sources.use = 18, targets.use = c(6:11),  
                 comparison = c(1, 2), angle.x = 45)
``` 


# 2. identify the up-regulated ligand-receptor pairs 
``` {r, fig.width= 12, fig.height= 15, message=FALSE, warning=FALSE}
netVisual_bubble(cellchat.compare, sources.use = 4, targets.use = c(6:11),  
                 comparison = c(1, 2), max.dataset = 2, 
                 title.name = "Increased signaling in LS", angle.x = 45, 
                 remove.isolate = T)
``` 
``` {r, fig.width= 12, fig.height= 15, message=FALSE, warning=FALSE}
# 3. identify down-regulated ligand-receptor pairs 
netVisual_bubble(cellchat.compare, sources.use = 4, targets.use = c(6:11),  
                 comparison = c(1, 2), max.dataset = 1, 
                 title.name = "Decreased signaling in LS", angle.x = 45, 
                 remove.isolate = T) 
``` 
``` {r, fig.width= 12, fig.height= 15, message=FALSE, warning=FALSE}
# Violin plot
View(cellchat@meta)
plotGeneExpression(cellchat.compare, signaling = "CXCL", split.by = "type", 
                   colors.ggplot = T)


saveRDS(cellchat, file = "cellchat.rds")
saveRDS(cellchat.first.art, file = "cellchat.first.rds")
saveRDS(cellchat.second.art, file = "cellchat.second.rds")
``` 
``` {r}
sessionInfo()
```
