---
title: "Clustering"
date: "2023-11-17"
output:
  html_document:
    toc: yes
    toc_depth: 2
    number_sections: yes
    toc_float:
      collapsed: yes
---

# Objective 

1- Performing extra QC by regressing out cell cycle phase S and G2/M and normalize using SCTransform

2- Integration with harmony and clustering 

# library
``` {r, warning= FALSE, fig.width= 6, fig.height= 6}
library(tidyverse)
library(Seurat)
library(patchwork)
library(sctransform)
library(glmGamPoi)
```

``` {r}
#==================================================================================================
#library(remotes)
# Replace 'X.X.X' with your desired version
install.packages("Matrix")
remotes::install_version(package = 'Seurat', version = package_version('4.0.3'))
remove.packages("Seurat")
#==================================================================================================
``` 

#load invidual objects
``` {r}
p_164S<- readRDS("Singlet.164S.RDS")
p_164S2<- readRDS("Singlet.164S2.RDS")
p_184S2<- readRDS("Singlet.184S2.RDS")
p_184S3<- readRDS("Singlet.184S3.RDS")
p_218S<- readRDS("Singlet.218S.RDS")
p_218S2<- readRDS("Singlet.218S2.RDS")
p_SA5<- readRDS("Singlet.SA5.RDS")
p_SA6<- readRDS("Singlet.SA6.RDS")
p_SA7<- readRDS("Singlet.SA7.RDS")

#remove unused objects to reduce memory
rm(p_164S, p_164S2, p_184S2, p_184S3, p_218S, p_218S2, p_SA5, p_SA6, p_SA7)
view(p_SA7@meta.data)

#merge seurat objects
Arthritis.merge<- merge(p_164S, y=c(p_164S2, p_184S2, p_184S3, p_218S, p_218S2, p_SA5, p_SA6, p_SA7))

#save merge object
saveRDS(Arthritis.merge, file = "Arthritis.merge.SCT.RDS") # save on 12/01/2022

#Remove merged object
rm(Arthritis.merge.flt)
``` 

# Read merged seurat object and plot QC plot

``` {r, fig.with= 8, fig.height= 6}
Arthritis.merge<- readRDS("Arthritis.merge.SCT.RDS")

#plot filtered data 
VlnPlot(Arthritis.merge, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
``` 


# Feature-Feature relationship
``` {r, fig.width= 10, fig.height= 5}
#Visualize feature-feature relationship 
plot1 <- FeatureScatter(Arthritis.merge, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident") + 
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) 
plot2 <- FeatureScatter(Arthritis.merge, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident") +
 theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
plot1|plot2 
``` 

#Filter noise genes

``` {r, fig.height= 6, fig.width= 7}
par(mar = c(4, 8, 2, 1))
C_u <- Arthritis.merge@assays$RNA@counts
C_u <- Matrix::t(Matrix::t(C_u)/Matrix::colSums(C_u)) * 100
most_expressed <- order(apply(C_u, 1, median), decreasing = T)[20:1]

#plot gene 
boxplot(as.matrix(t(C_u[most_expressed, ])), cex.axis = 0.5, las = 1, xlab = "% total count per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE)
``` 
# Filter noise genes 
``` {r}
# Filter MALAT1
Arthritis.merge.flt <- Arthritis.merge[!grepl("MALAT1", rownames(Arthritis.merge)), ]

# Filter Mitocondria
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^MT-", rownames(Arthritis.merge.flt)), ]

#Mitochondria
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^MT\\.", rownames(Arthritis.merge.flt)), ]

#Filter gene LOC
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^LOC(0-9)", rownames(Arthritis.merge.flt)), ]

# Filter Ribossomal gene 
Arthritis.merge.flt <- Arthritis.merge.flt[ ! grepl('^RP[SL]', rownames(Arthritis.merge.flt)), ]

# HSPA
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^HSPA", rownames(Arthritis.merge.flt)), ]

# MTRNR
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^MTRNR", rownames(Arthritis.merge.flt)), ]

# TR pattern
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^TR(A|B|G|D)V", rownames(Arthritis.merge.flt)), ]
```  

#Verify whether noise genes have been removed 

``` {r, fig.height= 6, fig.width= 7}
par(mar = c(4, 8, 2, 1))
C <- Arthritis.merge.flt@assays$RNA@counts
C <- Matrix::t(Matrix::t(C)/Matrix::colSums(C)) * 100
most_expressed <- order(apply(C, 1, median), decreasing = T)[20:1]

#plot gene 
boxplot(as.matrix(t(C[most_expressed, ])), cex.axis = 0.5, las = 1, xlab = "% total count per cell",
    col = (scales::hue_pal())(20)[20:1], horizontal = TRUE)
``` 

# SCTransform and regress out cell cycle score

``` {r, fig.width= 10, fig.height= 3 , warning= FALSE}

#First round of SCT
Arthritis<- SCTransform(Arthritis.merge.flt, 
                       assay = "RNA", 
                        new.assay.name = "SCT" , 
                       vars.to.regress = c("percent.mt", "nFeature_RNA", "nCount_RNA"), 
                       verbose = FALSE)

#Cell cycle score 
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

Arthritis<- CellCycleScoring(
  Arthritis,
  s.features = s.genes,
  g2m.features = g2m.genes,
  assay = 'SCT',
  set.ident = TRUE
  )

#plot cell cyle 
VlnPlot(Arthritis, features = c("S.Score", "G2M.Score"), group.by = "orig.ident",
    ncol = 4, pt.size = 0.1)

```

```{r}
# normalise again but this time including also the cell cycle scores
Arthritis<- SCTransform(
Arthritis,
assay = 'RNA',
new.assay.name = 'SCT',
vars.to.regress = c('percent.mt', 'nFeature_RNA', 'nCount_RNA', 'S.Score', 'G2M.Score')
)

# Save this object 
saveRDS(Arthritis, file = "Arthritis_SCT.RDS") #save on 12/01/23

# PCA
Arthritis <- RunPCA(Arthritis, verbose = FALSE, assay = "SCT", npcs = 50)

#Save this object
saveRDS(Arthritis, file="Arthritis_pca.RDS") # Saved on 12/01/2023

#Plot PC components ! 
Arthritis<- readRDS("Arthritis_pca.RDS")

#PC component plot
ElbowPlot(Arthritis, ndims = 50,reduction = "pca") +
  ggtitle("Number of principle component")  + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.title = element_text(hjust = 0.5))
```

#heatmap of highly variable genes 
```{r, fig.width= 8, fig.height= 5}
DimHeatmap(Arthritis, dims = 1:6, cells = 500, ncol = 3, balanced = TRUE)
``` 

# Integration using Harmony

``` {r, fig.width= 5, fig.height= 4}

#run Harmony integration
library(harmony)

#run the harmony function
Arthritis.Harmony<- RunHarmony(Arthritis, group.by.vars = c("orig.ident", "type"))

#save harmony object
#saveRDS(Arthritis.Harmony, file = "Harmony.SCT.RDS")

#RunUMAP
Arthritis.integration<- FindNeighbors(object = Arthritis.Harmony, 
                                      reduction="harmony", 
                                      dims = 1:43)

Arthritis.integration<- FindClusters(object = Arthritis.integration, 
                                     resolution = 0.4)

Arthritis.integration<- RunUMAP(object = Arthritis.integration, 
                                reduction = "harmony", 
                                dims = 1:43)


#Save final object
#saveRDS(Arthritis.integration, file= "UMAP.SCT.RDS")

#Read UMAP 
Arthritis.integration<- readRDS("UMAP.SCT.RDS")

#plot non-linear dimentionality reduction
DimPlot(Arthritis.integration, reduction = "umap",label = TRUE) +
  theme_minimal() +
  NoLegend()
```
# some feature plots

## CD4/Foxp3
``` {r, fig.width= 9, fig.height= 4}
FeaturePlot(Arthritis.integration, features = c("FOXP3", "CD4"), pt.size = 0.2, ncol = 2)
```
## T cell subsets
``` {r, fig.width= 7, fig.height= 5}

FeaturePlot(Arthritis.integration, features = c("CD3E", "CD8A", "CD4", "FOXP3"), pt.size = 0.2,
    ncol = 2)
```



## Some common markers 

``` {r, fig.width= 8, fig.height= 11}
FeaturePlot(Arthritis.integration, features = c("EPCAM", "PTPRC", "COL1A1", "CD79A", "CD3D", "S100A8", "S100A9", "CD8A"), pt.size = 0.2,
    ncol = 2)
```

``` {r, fig.width= 5, fig.height= 4}
Arthritis.integration.tsne<- RunTSNE(object = Arthritis.integration, 
                                reduction = "harmony", 
                                dims = 1:40)
DimPlot(Arthritis.integration.tsne, reduction = "tsne",label = TRUE) +
  theme_void() +
  NoLegend()
```



``` {r}
sessionInfo()
```