---
title: "Clustering"
date: "2023-11-15"
output:
  html_document:
    toc: yes
    toc_depth: 2
    number_sections: yes
    toc_float:
      collapsed: yes
---

# Objective 

1- Performing extra QC by regressing out cell cycle phase S and G2/M and normalize using SCTransform

2- Integration with harmony and clustering 

# library
``` {r, warning= FALSE, fig.width= 6, fig.height= 6}
library(tidyverse)
library(Seurat)
library(patchwork)
library(sctransform)
library(glmGamPoi)
```

``` {r}
#==================================================================================================
library(remotes)
# Replace 'X.X.X' with your desired version
remotes::install_version(package = 'Seurat', version = package_version('4.0.3'))
#==================================================================================================

``` {r}
p_164S<- readRDS("Singlet.164S.RDS")
p_164S2<- readRDS("Singlet.164S2.RDS")
p_184S2<- readRDS("Singlet.184S2.RDS")
p_184S3<- readRDS("Singlet.184S3.RDS")
p_218S<- readRDS("Singlet.218S.RDS")
p_218S2<- readRDS("Singlet.218S2.RDS")
p_SA5<- readRDS("Singlet.SA5.RDS")
p_SA6<- readRDS("Singlet.SA6.RDS")
p_SA7<- readRDS("Singlet.SA7.RDS")

#remove unsed object to reduce memory
rm(p_164S, p_164S2, p_184S2, p_184S3, p_218S, p_218S2, p_SA5, p_SA6, p_SA7)

#merge seurat objects
Arthritis.merge<- merge(p_164S, y=c(p_164S2, p_184S2, p_184S3, p_218S, p_218S2, p_SA5, p_SA6, p_SA7))

#Remove merged object
rm(Arthritis.merge)
``` 

# Read merged seurat object

``` {r}
Arthritis.merge<- readRDS("Arthritis.merge.SC.RDS")
``` 

# Feature-Feature relationship

``` {r, fig.width= 10, fig.height= 5}
#Visualize feature-feature relationship 
plot1 <- FeatureScatter(Arthritis.merge, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident") + 
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) 
plot2 <- FeatureScatter(Arthritis.merge, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident") +
 theme_minimal() + theme(plot.title = element_text(hjust = 0.5))

#comined plots 
pp<- plot1|plot2 
```


# SCTransform and regress out cell cycle score

``` {r, fig.width= 7, fig.height= 4, warning= FALSE}

#First round of SCT
Arthritis<- SCTransform(Arthritis.merge, 
                        assay = "RNA", 
                        new.assay.name = "SCT" , 
                        vars.to.regress = c("percent.mt", "nFeature_RNA", "nCount_RNA"), 
                        verbose = FALSE)

#Cell cycle score 
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

Arthritis<- CellCycleScoring(
  Arthritis,
  s.features = s.genes,
  g2m.features = g2m.genes,
  assay = 'SCT',
  set.ident = TRUE
  )

# normalise again but this time including also the cell cycle scores
Arthritis<- SCTransform(
  Arthritis,
  assay = 'RNA',
  new.assay.name = 'SCT',
  vars.to.regress = c('percent.mt', 'nFeature_RNA', 'nCount_RNA', 'S.Score', 'G2M.Score')
  )

# PCA
Arthritis <- RunPCA(Arthritis, seed.use = 12345,verbose = FALSE, assay = "SCT", npcs = 50)


#Plot PC components ! 
ElbowPlot(Arthritis, ndims = 50,reduction = "pca") +
  ggtitle("Number of principle component")  + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

# Integration using Harmony

``` {r, fig.width= 5, fig.height= 4}
library(harmony)

#run Harmony integration
Arthritis.integration<- RunHarmony(Arthritis, group.by.vars = "orig.ident")
Arthritis.integration<- RunUMAP(object = Arthritis.integration, 
                                reduction = "harmony", 
                                dims = 1:40, 
                                min.dist = 0.2, spread = 2,
                                seed.use = 12345)
Arthritis.integration<- FindNeighbors(object = Arthritis.integration, reduction="harmony", dims = 1:40)
Arthritis.integration<- FindClusters(object = Arthritis.integration, resolution = 0.6, random.seed = 12345)

#plot non-linear dimentionality reduction
DimPlot(Arthritis.integration, label = TRUE) +
  theme_void() +
  NoLegend()
```

# some feature plots
## T cell subsets
``` {r, fig.width= 7, fig.height= 5}

FeaturePlot(Arthritis.integration, features = c("CD3E", "CD8A", "CD4", "FOXP3"), pt.size = 0.2,
    ncol = 2)
```

## CD4/Foxp3
``` {r, fig.width= 9, fig.height= 4}
FeaturePlot(Arthritis.integration, features = c("CD4", "FOXP3"), pt.size = 0.2,
    ncol = 2)
```

## Some common markers 

``` {r, fig.width= 8, fig.height= 9}
FeaturePlot(Arthritis.integration, features = c("EPCAM", "PTPRC", "COL1A1", "CD79A", "CD3D", "S100A8", "S100A9", "CD8A"), pt.size = 0.2,
    ncol = 2)

```

``` {r}
sessionInfo()
```



``` {r}
sessionInfo()
```





