---
title: "SCTransform_part2"
output: html_document
date: "2023-11-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Integration using harmony
``` {r}
library(harmony)
set.seed(12345)
Arthritis<- RunHarmony(Arthritis ,group.by.vars = "orig.ident")
```

# Find Neighbors and clusters
``` {r, fig.width= 8, fig.height= 10}
set.seed(12345)
Arthritis<- FindNeighbors(Arthritis, reduction = "harmony", dims = 1:30, verbose = FALSE)

#run resolution between 0 and 1
resolution_values <- seq(0, 1, by = 0.1)
Arthritis.clustree<- FindClusters(Arthritis, resolution = resolution_values, random.seed = 12345, verbose = FALSE)
``` 

# examine cluster tree to deterine the optimal resolution 

``` {r, fig.width= 8, fig.height= 9}
library(clustree)
clustree(Arthritis.clustree)
```
# Save cluster tree object

``` {r}
saveRDS(Arthritis, file ="Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/Cluster.tree.RDS")
``` 

# Investigate various min.dist value of RunUMAP function
``` {r}
# Determine the min.dist values you want to test
min_dist_values <- c(0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1, 0.05, 0.01, 0.005, 0.001)  

# Create a list to store the plots
plot_list <- list()

for(dist in min_dist_values){
    Arthritis <- RunUMAP(object = Arthritis ,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = dist)

    Arthritis <- FindNeighbors(object = Arthritis, reduction="harmony", dims = 1:30)
    snn.obj <- FindClusters(object = Arthritis, resolution = 0.2, random.seed = 12345)
    
    umap_plot <- DimPlot(snn.obj, reduction="umap", ncol=2, group.by = "seurat_clusters") +
      ggtitle(paste("UMAP Plot (min.dist =", dist, ")"))
    
    # Append the plot to the list
    plot_list[[as.character(dist)]] <- umap_plot
}
```

# Combine all plots into a single figure.
``` {r}
library(ggplot2)
library(gridExtra)

# Calculate the number of rows and columns for the grid
total_plots <- length(min_dist_values)
n_rows <- ceiling(sqrt(total_plots))
n_cols <- ceiling(total_plots / n_rows)

# Combine all plots into a flexible grid
combined_plot <- do.call(grid.arrange, c(plot_list, ncol = n_cols, nrow = n_rows))

# Save the combined plot
ggsave("combined_min_dist_plots.png", combined_plot, height=18, width=20)
``` 

# Or Save individual plot of min.dist separately  
``` {r}
# Create a directory to save the plots 
dir.create("min_dist_plots", showWarnings = FALSE)

#iterate through various min.dist values
for(dist in c(0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1, 0.05, 0.01, 0.005, 0.001)){
    Arthritis.dist <- RunUMAP(object = Arthritis,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = dist)

    Arthritis.dist <- FindNeighbors(object = Arthritis.dist, reduction="harmony", dims = 1:30)
    snn.obj <- FindClusters(object = Arthritis.dist, resolution = 0.2)
    
    umap_plot <- DimPlot(snn.obj, reduction="umap", ncol=2, group.by = "seurat_clusters", label = TRUE) +
      NoLegend() +
      ggtitle(paste("UMAP Plot (min.dist =", dist, ")"))
    
    
    #Save the plot
    plot_filename <- paste0("min_dist_plots/UMAP_min_dist_", dist, ".png")
    ggsave(filename = plot_filename, plot = umap_plot, height=4, width=7, dpi=300)
}
``` 

# Examine various resolution values
``` {r}
# Create a directory to save the plots 
dir.create("Resolution_plots", showWarnings = FALSE)
# Run UMAP
Arthritis.res<- RunUMAP(object = Arthritis,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = 0.01)
Arthritis.res <- FindNeighbors(object =  Arthritis.res, reduction="harmony", dims = 1:30)

#iterate through various value of resolutions 
for(res in c(0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)){
    snn.obj.res <- FindClusters(object = Arthritis.res, resolution = res, random.seed = 12345, verbose = FALSE)
    
    umap_plot_res <- DimPlot(snn.obj.res, reduction="umap", ncol=2, group.by = "seurat_clusters", label = TRUE) +
      NoLegend() +
      ggtitle(paste("UMAP Plot (resolution =", res, ")"))
    
    #Save the plot
    plot_filename_res <- paste0("Resolution_plots/UMAP_resoltuion_", res, ".png")
    ggsave(filename = plot_filename_res, plot = umap_plot_res, height=4, width=7, dpi=300)
    
}
``` 

# Testing various value of spread of RunUMAP function

``` {r}
# Create a directory to save the plots 
dir.create("spread_plots", showWarnings = FALSE)

#iterate through various min.dist values
for(spr in c(0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4)){
    Arthritis.spr <- RunUMAP(object = Arthritis,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = 0.01, spread = spr)

    Arthritis.spr <- FindNeighbors(object = Arthritis.spr, reduction="harmony", dims = 1:30)
    snn.obj.spr <- FindClusters(object = Arthritis.spr, resolution = 0.6)
    
    umap_plot_spr <- DimPlot(snn.obj.spr, reduction="umap", ncol=2, group.by = "seurat_clusters", label = TRUE) +
      NoLegend() +
      ggtitle(paste("UMAP Plot (spread =", spr, ")"))
    
    
    #Save the plot
    plot_filename_spr <- paste0("spread_plots/UMAP_spread_", spr, ".png")
    ggsave(filename = plot_filename_spr, plot = umap_plot_spr, height=4, width=7, dpi=300)
}
``` 


# Findning optimal number of n.neighbor of RunUMAP

``` {r}
# Create a directory to save the plots 
dir.create("neighbors_plots", showWarnings = FALSE)

#iterate through various min.dist values
for(neigh in c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50)){
    Arthritis.neigh <- RunUMAP(object = Arthritis,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = 0.1, spread = 3.5, n.neighbors = neigh)

    Arthritis.neigh  <- FindNeighbors(object = Arthritis.neigh , reduction="harmony", dims = 1:30)
    snn.obj.neigh <- FindClusters(object = Arthritis.neigh, resolution = 0.6, random.seed = 12345)
    
    umap_plot_neigh <- DimPlot(snn.obj.neigh, reduction="umap", ncol=2, group.by = "seurat_clusters", label = TRUE) +
      NoLegend() +
      ggtitle(paste("UMAP Plot (n.neighbors =", neigh, ")"))
    
    
    #Save the plot
    plot_filename_neigh <- paste0("neighbors_plots/neighbor", neigh, ".png")
    ggsave(filename = plot_filename_neigh , plot = umap_plot_neigh, height=4, width=7, dpi=300)
}
``` 

# Run single UMAP using optimal value found in previous section. 
``` {r}
#read normalize files back in 
Arthritis.singleparam<- readRDS("SCTranform.RDS")

#Run PCA
Arthritis.singleparam<- RunPCA(Arthritis.singleparam,seed.use = 12345, verbose = FALSE)

#Elbow plot
ElbowPlot(Arthritis.singleparam, ndims = 50,reduction = "pca") +
  ggtitle("Number of principle component")  + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

#Run single UMAP with optimal parameter!

``` {r, fig.width= 5, fig.height= 5}
library(harmony)
set.seed(12345)

#integration
UMAP.singlparam<- RunHarmony(Arthritis.singleparam ,group.by.vars = "orig.ident")

#RUNUMAP pipeline 
UMAP.singlparam<- RunUMAP(object = UMAP.singlparam,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = 0.2, spread = 3, n.neighbors = 30)
UMAP.singlparam<- FindNeighbors(UMAP.singlparam, reduction = "harmony", dims = 1:30, verbose = FALSE)
UMAP.singlparam<- FindClusters(UMAP.singlparam, resolution = 0.6, random.seed = 12345)

#plot umap 
DimPlot(UMAP.singlparam) +
  theme_minimal() +
  NoLegend()
``` 






