---
title: "Single cell Seurat/SCTransform"
date: "2023-11-08"
output:
  html_document:
    toc: yes
    toc_depth: 2
    number_sections: yes
    toc_float:
      collapsed: yes
---

# Objective

1. Perform standard preprocessing steps
2. Integration using Harmony
3. Clustering and trick for fine-tune UMPA plots (finding best UMAP plots)

# Load the required library 
``` {r, warning= FALSE}
library(tidyverse)
library(Seurat)
library(patchwork)
library(sctransform)
library(glmGamPoi)
```

# Load individual data of each samples
``` {r}
#164S !
A164S<- Read10X(data.dir = "164S")
A164S2<- Read10X(data.dir = "164S2")

#184S !
A184S2<- Read10X(data.dir = "184S2")
A184S3<- Read10X(data.dir = "184S3")

#218S !
A218S<- Read10X(data.dir = "218S")
A218S2<- Read10X(data.dir = "218S2")

#SA !
ASA5<- Read10X(data.dir = "SA5")
ASA6<- Read10X(data.dir = "SA6")
ASA7<- Read10X(data.dir = "SA7")
```

# Create Seurat Object
```{r, echo=FALSE}
#164S !
A164S<- CreateSeuratObject(counts = A164S, project = "164S", min.cells = 3, min.features = 200)

#164S2 !
A164S2<- CreateSeuratObject(counts = A164S2, project = "164S2", min.cells = 3, min.features = 200)

#184S2 !
A184S2<- CreateSeuratObject(counts = A184S2, project = "184S2", min.cells = 3, min.features = 200)

#184S3 !
A184S3<- CreateSeuratObject(counts = A184S3, project = "184S3", min.cells = 3, min.features = 200)

#218S !
A218S<- CreateSeuratObject(counts = A218S, project = "218S", min.cells = 3, min.features = 200)

#A218S2 ! 
A218S2<- CreateSeuratObject(counts = A218S2, project = "218S2", min.cells = 3, min.features = 200)

#SA5 !
ASA5<- CreateSeuratObject(counts = ASA5, project = "SA5", min.cells = 3, min.features = 200)

#ASA6 !
ASA6<- CreateSeuratObject(counts = ASA6, project = "SA6", min.cells = 3, min.features = 200)

#SA7 !
ASA7<- CreateSeuratObject(counts = ASA7, project = "SA7", min.cells = 3, min.features = 200)
```

# Adding type variable to seurat object (e.g., First arthritis, recurrent arthritis and Osteoarthritis)
``` {r}
A164S$type= "First arthritis"  
A184S2$type = "First arthritis" 
A218S$type = "First arthritis"  
A164S2$type= "Second arthritis"  
A184S3$type = "Second arthritis" 
A218S2$type = "Second arthritis" 
ASA5$type = "Osteoarthritis"
ASA6$type = "Osteoarthritis"
ASA7$type = "Osteoarthritis"
``` 

# Merge Seurat Object and plot unfiltered cells
``` {r, fig.width= 9, fig.height= 5}
#merge objects
Arthritis.merge<- merge(A164S, y=c(A164S2, A184S2, A184S3, A218S, A218S2, ASA5, ASA6, ASA7))

#set percent MT
Arthritis.merge<- PercentageFeatureSet(Arthritis.merge, pattern = "^MT-", col.name = "percent.mt")

#plot unfiltered data 
unfiltered<- VlnPlot(Arthritis.merge, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
unfiltered
```


# Filter low quality cells based on nFeatureRNA, nCountRNA and % Mitochondria genes
``` {r, fig.width= 9, fig.height= 5}
#Set the bar!
Filtered.cells <- subset(Arthritis.merge, subset = nFeature_RNA >500 
              & nFeature_RNA <5500 & nCount_RNA>800 & nCount_RNA <31000 & percent.mt < 15)

#save object
saveRDS(Filtered.cells, file = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/Filtered_cells.RDS")


#plot filtered cells!  
Filtered<- VlnPlot(Filtered.cells, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
Filtered
``` 

#plot unfiltered & filtered cells in paralell
``` {r, fig.width= 9, fig.height= 10}
library(patchwork)
plot1<- unfiltered/Filtered
plot1

#

#save high quality plot 
ggsave("plot1.tiff" , plot = plot1, height=8, width = 9, units = "in",  dpi=300)
```


# Filter noise genes 
``` {r}
# Filter MALAT1
Arthritis.merge.flt <- Filtered.cells[!grepl("MALAT1", rownames(Filtered.cells)), ]

# Filter Mitocondria
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^MT-", rownames(Arthritis.merge.flt)), ]

#Mitochondria
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^MT\\.", rownames(Arthritis.merge.flt)), ]

#Filter gene LOC
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^LOC(0-9)", rownames(Arthritis.merge.flt)), ]

# Filter Ribossomal gene 
Arthritis.merge.flt <- Arthritis.merge.flt[ ! grepl('^RP[SL]', rownames(Arthritis.merge.flt)), ]

# HSPA
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^HSPA", rownames(Arthritis.merge.flt)), ]

# MTRNR
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^MTRNR", rownames(Arthritis.merge.flt)), ]

# TR pattern
Arthritis.merge.flt <- Arthritis.merge.flt[!grepl("^TR(A|B|G|D)V", rownames(Arthritis.merge.flt)), ]
```  

# Normalize cell using SCTransform

``` {r}
#SCTransform
Arthritis<- SCTransform(Arthritis.merge.flt, 
                        method= "glmGamPoi", 
                        vars.to.regress = "percent.mt", 
                        verbose= FALSE)

#save SCTransform object
saveRDS(Arthritis, file = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/SCTranform.RDS")

#read normalize SCTransform back in 
Arthritis<- readRDS("SCTranform.RDS")
``` 


# plot Feature-Feature relationship
``` {r, fig.width= 10, fig.height= 5}
#Visualize feature-feature relationship 
#
plot2 <- FeatureScatter(Arthritis, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident") + 
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) 
plot3 <- FeatureScatter(Arthritis, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident") + theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
plot2|plot3 
```



# Perform dimensionality reduction by PCA and plot PC components
``` {r}

#Run PCA
Arthritis<- RunPCA(Arthritis, verbose = FALSE)

#Elbow plot
ElbowPlot(Arthritis, ndims = 50,reduction = "pca") +
  ggtitle("Number of principle component")  + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
``` 

# Heatmap for Deg

``` {r, fig.width= 8, fig.height= 5}
DimHeatmap(Arthritis, dims = 1:3, cells = 500, balanced = TRUE)
``` 

# Integration using harmony
``` {r}
library(harmony)
set.seed(12345)
Arthritis<- RunHarmony(Arthritis ,group.by.vars = "orig.ident")
```

# Find Neighbors and clusters
``` {r, fig.width= 8, fig.height= 10}
set.seed(12345)
Arthritis<- FindNeighbors(Arthritis, reduction = "harmony", dims = 1:30, verbose = FALSE)

#run resolution between 0 and 1
resolution_values <- seq(0, 1, by = 0.1)
Arthritis.clustree<- FindClusters(Arthritis, resolution = resolution_values, random.seed = 12345, verbose = FALSE)
``` 

# examine cluster tree to deterine the optimal resolution 

``` {r, fig.width= 8, fig.height= 9}
library(clustree)
clustree(Arthritis.clustree)
```
# Save cluster tree object

``` {r}
saveRDS(Arthritis, file ="Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/Cluster.tree.RDS")
``` 

# Investigate various min.dist value of RunUMAP function

``` {r}
# Determine the min.dist values you want to test
min_dist_values <- c(0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1, 0.05, 0.01, 0.005, 0.001)  

# Create a list to store the plots
plot_list <- list()

for(dist in min_dist_values){
    Arthritis <- RunUMAP(object = Arthritis ,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = dist)

    Arthritis <- FindNeighbors(object = Arthritis, reduction="harmony", dims = 1:30)
    snn.obj <- FindClusters(object = Arthritis, resolution = 0.2, random.seed = 12345)
    
    umap_plot <- DimPlot(snn.obj, reduction="umap", ncol=2, group.by = "seurat_clusters") +
      ggtitle(paste("UMAP Plot (min.dist =", dist, ")"))
    
    # Append the plot to the list
    plot_list[[as.character(dist)]] <- umap_plot
}
```

## Combine all plots into a single figure.
``` {r}
library(ggplot2)
library(gridExtra)

# Calculate the number of rows and columns for the grid
total_plots <- length(min_dist_values)
n_rows <- ceiling(sqrt(total_plots))
n_cols <- ceiling(total_plots / n_rows)

# Combine all plots into a flexible grid
combined_plot <- do.call(grid.arrange, c(plot_list, ncol = n_cols, nrow = n_rows))

# Save the combined plot
ggsave("combined_min_dist_plots.png", combined_plot, height=18, width=20)
``` 

Comments: I saved these combined min.dist plots and can be viewed outside R Markdown. 


## Or Save individual plot of min.dist separately  
``` {r}
# Create a directory to save the plots 
dir.create("min_dist_plots", showWarnings = FALSE)

#iterate through various min.dist values
for(dist in c(0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1, 0.05, 0.01, 0.005, 0.001)){
    Arthritis.dist <- RunUMAP(object = Arthritis,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = dist)

    Arthritis.dist <- FindNeighbors(object = Arthritis.dist, reduction="harmony", dims = 1:30)
    snn.obj.dist <- FindClusters(object = Arthritis.dist, resolution = 0.2)
    
    umap_plot.dist <- DimPlot(snn.obj.dist, reduction="umap", group.by = "seurat_clusters", label = TRUE) +
      NoLegend() +
      ggtitle(paste("UMAP Plot (min.dist =", dist, ")"))
    
    
    #Save the plot
    plot_filename <- paste0("min_dist_plots/UMAP_min_dist_", dist, ".png")
    ggsave(filename = plot_filename, plot = umap_plot.dist, height=4, width=5, dpi=300)
}
``` 

# Examine various resolution values
``` {r}
# Create a directory to save the plots 
dir.create("Resolution_plots", showWarnings = FALSE)
# Run UMAP
Arthritis.res<- RunUMAP(object = Arthritis,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = 0.01)
Arthritis.res <- FindNeighbors(object =  Arthritis.res, reduction="harmony", dims = 1:30)

#iterate through various value of resolutions 
for(res in c(0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)){
    snn.obj.res <- FindClusters(object = Arthritis.res, resolution = res, random.seed = 12345, verbose = FALSE)
    
    umap_plot_res <- DimPlot(snn.obj.res, reduction="umap", ncol=2, group.by = "seurat_clusters", label = TRUE) +
      NoLegend() +
      ggtitle(paste("UMAP Plot (resolution =", res, ")"))
    
    #Save the plot
    plot_filename_res <- paste0("Resolution_plots/UMAP_resoltuion_", res, ".png")
    ggsave(filename = plot_filename_res, plot = umap_plot_res, height=4, width=7, dpi=300)
    
}
``` 

# Testing various value of spread of RunUMAP function

``` {r}
# Create a directory to save the plots 
dir.create("spread_plots", showWarnings = FALSE)

#iterate through various min.dist values
for(spr in c(0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4)){
    Arthritis.spr <- RunUMAP(object = Arthritis,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = 0.01, spread = spr)

    Arthritis.spr <- FindNeighbors(object = Arthritis.spr, reduction="harmony", dims = 1:30)
    snn.obj.spr <- FindClusters(object = Arthritis.spr, resolution = 0.6)
    
    umap_plot_spr <- DimPlot(snn.obj.spr, reduction="umap", ncol=2, group.by = "seurat_clusters", label = TRUE) +
      NoLegend() +
      ggtitle(paste("UMAP Plot (spread =", spr, ")"))
    
    
    #Save the plot
    plot_filename_spr <- paste0("spread_plots/UMAP_spread_", spr, ".png")
    ggsave(filename = plot_filename_spr, plot = umap_plot_spr, height=4, width=7, dpi=300)
}
``` 


# Findning optimal number of n.neighbor of RunUMAP

``` {r}
# Create a directory to save the plots 
dir.create("neighbors_plots", showWarnings = FALSE)

#iterate through various min.dist values
for(neigh in c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50)){
    Arthritis.neigh <- RunUMAP(object = Arthritis,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = 0.1, spread = 3.5, n.neighbors = neigh)

    Arthritis.neigh  <- FindNeighbors(object = Arthritis.neigh , reduction="harmony", dims = 1:30)
    snn.obj.neigh <- FindClusters(object = Arthritis.neigh, resolution = 0.6, random.seed = 12345)
    
    umap_plot_neigh <- DimPlot(snn.obj.neigh, reduction="umap", ncol=2, group.by = "seurat_clusters", label = TRUE) +
      NoLegend() +
      ggtitle(paste("UMAP Plot (n.neighbors =", neigh, ")"))
    
    
    #Save the plot
    plot_filename_neigh <- paste0("neighbors_plots/neighbor", neigh, ".png")
    ggsave(filename = plot_filename_neigh , plot = umap_plot_neigh, height=4, width=7, dpi=300)
}
``` 

# Run single UMAP using optimal value found in previous section. 
``` {r}
#read normalize files back in 
Arthritis.singleparam<- readRDS("SCTranform.RDS")

#Run PCA
Arthritis.singleparam<- RunPCA(Arthritis.singleparam,seed.use = 12345, verbose = FALSE)

#Elbow plot
ElbowPlot(Arthritis.singleparam, ndims = 50,reduction = "pca") +
  ggtitle("Number of principle component")  + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

#Run single UMAP with optimal parameter!

``` {r, fig.width= 6, fig.height= 4}
library(harmony)
set.seed(12345)

#integration
UMAP.singlparam<- RunHarmony(Arthritis,group.by.vars = "orig.ident")

#RUNUMAP pipeline 
UMAP.singlparam<- FindNeighbors(UMAP.singlparam, reduction = "harmony", dims = 1:40, verbose = FALSE)
UMAP.singlparam<- FindClusters(UMAP.singlparam, resolution = 0.4)
UMAP.singlparam<- RunUMAP(object = UMAP.singlparam,
                          reduction = "harmony",
                          dims = 1:40)


#plot umap 
DimPlot(UMAP.singlparam, reduction= "umap", label = TRUE, repel = FALSE
        ) +
  theme_minimal() +
  NoLegend()
``` 

## T cell subsets
``` {r, fig.width= 7, fig.height= 5}

FeaturePlot(UMAP.singlparam, features = c("CD3E", "CD8A", "CD4", "FOXP3"), pt.size = 0.2,
    ncol = 2)
```


``` {r}
sessionInfo()
```

















