---
title: "preprocessing and TSNE"
date: "2023-11-17"
output:
  html_document:
    toc: yes
    toc_depth: 2
    number_sections: yes
    toc_float:
      collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sincle cell RNASeq of recurrent arthritis part 1
## Objective

1. Perform reprocessing such as remove low quality cells 
2. Normalization, feature scaling and clustering using TSNE and UMAP
3. Examine cell cycling

# Load data and library
```{r}
library(tidyverse)
library(Seurat)
library(patchwork)
set.seed(198752)
```

# Read in individual data
``` {r}
#load individual data of each patient

#164S !
A164S<- Read10X(data.dir = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/164S/")
A164S2<- Read10X(data.dir = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/164S2/")

#184S !
A184S2<- Read10X(data.dir = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/184S2/")
A184S3<- Read10X(data.dir = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/184S3/")

#218S !
A218S<- Read10X(data.dir = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/218S/")
A218S2<- Read10X(data.dir = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/218S2/")

#SA !
ASA5<- Read10X(data.dir = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/SA5/")
ASA6<- Read10X(data.dir = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/SA6/")
ASA7<- Read10X(data.dir = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/SA7/")
```

# Create Seurat Objects
```{r}
#164S !
A164S<- CreateSeuratObject(counts = A164S, project = "164S", min.cells = 3, min.features = 200)

#164S2 !
A164S2<- CreateSeuratObject(counts = A164S2, project = "164S2", min.cells = 3, min.features = 200)

#184S2 !
A184S2<- CreateSeuratObject(counts = A184S2, project = "184S2", min.cells = 3, min.features = 200)

#184S3 !
A184S3<- CreateSeuratObject(counts = A184S3, project = "184S3", min.cells = 3, min.features = 200)

#218S !
A218S<- CreateSeuratObject(counts = A218S, project = "218S", min.cells = 3, min.features = 200)

#A218S2 ! 
A218S2<- CreateSeuratObject(counts = A218S2, project = "218S2", min.cells = 3, min.features = 200)

#SA5 !
ASA5<- CreateSeuratObject(counts = ASA5, project = "SA5", min.cells = 3, min.features = 200)

#ASA6 !
ASA6<- CreateSeuratObject(counts = ASA6, project = "SA6", min.cells = 3, min.features = 200)

#SA7 !
ASA7<- CreateSeuratObject(counts = ASA7, project = "SA7", min.cells = 3, min.features = 200)
```

# Divide samples to three groups
``` {r}
A164S$type= "First arthritis"  
A184S2$type = "First arthritis" 
A218S$type = "First arthritis"  
A164S2$type= "Second arthritis"  
A184S3$type = "Second arthritis" 
A218S2$type = "Second arthritis" 
ASA5$type = "Osteoarthritis"
ASA6$type = "Osteoarthritis"
ASA7$type = "Osteoarthritis"
``` 


# Filter low quality cells for each sample !
``` {r, fig.width= 7, fig.height= 6}
A164S <- PercentageFeatureSet(A164S, pattern = "^MT-", col.name = "percent.mt")
A164S_unfiltered<- VlnPlot(A164S, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#filtered low quality cells
A164S<- subset(A164S, subset = nFeature_RNA > 500 
              & nFeature_RNA < 6300 & nCount_RNA> 500 & nCount_RNA <30000 & percent.mt < 15)
A164S_filtered<- VlnPlot(A164S, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#plot unfiltered and filtered cells 
A164S.plot<- A164S_unfiltered/A164S_filtered
A164S.plot
```

``` {r, fig.width= 7, fig.height= 6}
A164S2 <- PercentageFeatureSet(A164S2, pattern = "^MT-", col.name = "percent.mt")
A164S2_unfiltered<- VlnPlot(A164S2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#filtered low quality cells
A164S2<- subset(A164S2, subset = nFeature_RNA > 500 
              & nFeature_RNA < 6500 & nCount_RNA> 800 & nCount_RNA <30000 & percent.mt < 15)
A164S2_filtered<- VlnPlot(A164S2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#plot unfiltered and filtered cells 
A164S2.plot<- A164S2_unfiltered/A164S2_filtered
A164S2.plot
```

``` {r, fig.width= 7, fig.height= 6}
A184S2 <- PercentageFeatureSet(A184S2, pattern = "^MT-", col.name = "percent.mt")
A184S2_unfiltered<- VlnPlot(A184S2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#filtered low quality cells
A184S2<- subset(A184S2, subset = nFeature_RNA > 500 
              & nFeature_RNA < 5800 & nCount_RNA> 500 & nCount_RNA <28000 & percent.mt < 15)
A184S2_filtered<- VlnPlot(A184S2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 

#plot unfiltered and filtered cells 
A184S2.plot<- A184S2_unfiltered/A184S2_filtered
A184S2.plot
```

``` {r, fig.width= 7, fig.height= 6}
A184S3 <- PercentageFeatureSet(A184S3, pattern = "^MT-", col.name = "percent.mt")
A184S3_unfiltered<- VlnPlot(A184S3, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#filtered low quality cells
A184S3<- subset(A184S3, subset = nFeature_RNA > 500 
              & nFeature_RNA < 6200 & nCount_RNA> 800 & nCount_RNA <34000 & percent.mt < 10)
A184S3_filtered<- VlnPlot(A184S3, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 

#plot unfiltered and filtered cells 
A184S3.plot<- A184S2_unfiltered/A184S2_filtered
A184S3.plot
```

``` {r, fig.width= 7, fig.height= 6}
A218S <- PercentageFeatureSet(A218S, pattern = "^MT-", col.name = "percent.mt")
A218S_unfiltered<- VlnPlot(A218S, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#filtered low quality cells
A218S<- subset(A218S, subset = nFeature_RNA > 500 
              & nFeature_RNA < 5800 & nCount_RNA> 500 & nCount_RNA <23000 & percent.mt < 15)
A218S_filtered<- VlnPlot(A218S, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 

#plot unfiltered and filtered cells 
A218S.plot<- A218S_unfiltered/A218S_filtered
A218S.plot
```

``` {r, fig.width= 7, fig.height= 6}
A218S2 <- PercentageFeatureSet(A218S2, pattern = "^MT-", col.name = "percent.mt")
A218S2_unfiltered<- VlnPlot(A218S2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#filtered low quality cells
A218S2<- subset(A218S2, subset = nFeature_RNA > 500 
              & nFeature_RNA < 6000 & nCount_RNA> 500 & nCount_RNA <25000 & percent.mt < 15)
A218S2_filtered<- VlnPlot(A218S2, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 

#plot unfiltered and filtered cells 
A218S2.plot<- A218S2_unfiltered/A218S2_filtered
A218S2.plot
```

``` {r, fig.width= 7, fig.height= 6}
ASA5 <- PercentageFeatureSet(ASA5, pattern = "^MT-", col.name = "percent.mt")
ASA5_unfiltered<- VlnPlot(ASA5, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#filtered low quality cells
ASA5<- subset(ASA5, subset = nFeature_RNA > 500 
              & nFeature_RNA < 7000 & nCount_RNA> 500& nCount_RNA <30000 & percent.mt < 15)
ASA5_filtered<- VlnPlot(ASA5, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 

#plot unfiltered and filtered cells 
ASA5.plot<- ASA5_unfiltered/ASA5_filtered
ASA5.plot
```

``` {r, fig.width= 7, fig.height= 6}
ASA6 <- PercentageFeatureSet(ASA6, pattern = "^MT-", col.name = "percent.mt")
ASA6_unfiltered<- VlnPlot(ASA6, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#filtered low quality cells
ASA6<- subset(ASA6, subset = nFeature_RNA > 500 
              & nFeature_RNA < 7500 & nCount_RNA> 800 & nCount_RNA <50000 & percent.mt < 15)
ASA6_filtered<- VlnPlot(ASA6, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 

#plot unfiltered and filtered cells 
ASA6.plot<- ASA6_unfiltered/ASA6_filtered
ASA6.plot
```

``` {r, fig.width= 7, fig.height= 6}
ASA7 <- PercentageFeatureSet(ASA7, pattern = "^MT-", col.name = "percent.mt")
ASA7_unfiltered<- VlnPlot(ASA7, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#filtered low quality cells
ASA7<- subset(ASA7, subset = nFeature_RNA > 500 
              & nFeature_RNA < 7500 & nCount_RNA> 800 & nCount_RNA <48000 & percent.mt < 10)
ASA7_filtered<- VlnPlot(ASA7, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3) 

#plot unfiltered and filtered cells 
ASA7.plot<- ASA7_unfiltered/ASA7_filtered
ASA7.plot
```

# Merge seurat objects
``` {r, fig.width= 10, fig.height= 5}
library(tidyverse)
#merge seurat object 
All_arthritis <- merge(A164S, y=c(A164S2, A184S2, A184S3, A218S, A218S2, ASA5, ASA6, ASA7))
#rm(All_arthritis)

#save this object
#saveRDS(All_arthritis, file = "Y:/All_arthritis.RDS")
``` 

# Feature-Feature relationship
``` {r, fig.width= 10, fig.height= 5}
#Visualize feature-feature relationship 
#
plot1 <- FeatureScatter(All_arthritis, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident") + 
  theme_minimal() + theme(plot.title = element_text(hjust = 0.5)) 
plot2 <- FeatureScatter(All_arthritis, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident") +
 theme_minimal() + theme(plot.title = element_text(hjust = 0.5))
plot1|plot2 
```

# see highly variable genes
``` {r, fig.width= 10, fig.height= 4}
# Seurat Standard work flow (SWF)
set.seed(198752)
All_arthritis_normal <- NormalizeData(All_arthritis) 
All_arthritis_normal <- FindVariableFeatures(All_arthritis_normal,
                                             selection.method = "vst",
                                             nfeatures = 2000, 
                                             verbose = FALSE) 
top10<- head(VariableFeatures(All_arthritis_normal), 10)
top10
```

#plot highly variable features!
``` {r, warning= FALSE, fig.height= 5, fig.width= 7}
High_feature1 <- VariableFeaturePlot(All_arthritis_normal)
High_feature2 <- LabelPoints(plot = High_feature1, points = top10, repel = TRUE)
High_feature2
```

# Determine the number of principle coponents that explain variation in the data
``` {r}
#feature scaling !
All_arthritis_normal <- ScaleData(All_arthritis_normal, 
                                  vars.to.regress = c("percent.mt", "nFeature_RNA", "nCount_RNA"),
                                  verbose = FALSE) 

#Run PCA !
All_arthritis_normal <- RunPCA(All_arthritis_normal, seed.use = 25483,verbose = FALSE)

#Plot PC components ! 
ElbowPlot(All_arthritis_normal, ndims = 50,reduction = "pca") +
  ggtitle("Number of principle component")  + 
  theme(plot.title = element_text(hjust = 0.5))
``` 

# Visulization of PC1 and PC2 in heatmap 
``` {r, fig.width= 8, fig.height= 7}
#heatmap of the first principle component
DimHeatmap(All_arthritis_normal, dims = 1:6, cells = 500, balanced = TRUE, ncol = 3)
```

# Integration using Harmony for TSNE
``` {r}
#Integration using harmony for tsne
library(harmony)
set.seed(198752)
All_arthritis_integration<- RunHarmony(All_arthritis_normal, group.by.vars = "orig.ident")

# save integration object
#saveRDS(All_arthritis_integration, file = "Y:/Recurrent__Arthritis__Single cell/Single_Cell_Arthritis/harmony.RDS")

# Run TSNE
All_arthritis_integration_TSNE<- RunTSNE(All_arthritis_integration, reduction = "harmony", dims = 1:40)
All_arthritis_integration_TSNE <- FindNeighbors(All_arthritis_integration_TSNE, reduction= "harmony", dims = 1:40)
All_arthritis_integration_TSNE <- FindClusters(All_arthritis_integration_TSNE, resolution = 0.4)

# Look at cluster IDs of the first 5 cells
head(Idents(All_arthritis_integration_TSNE), 5)
```

# Visualize all samples in TSNE 
``` {r, fig.width= 7, fig.height= 6}
#Run non-linear dimensional reduction
DimPlot(All_arthritis_integration_TSNE, reduction = "tsne", label = TRUE, size= 10) + 
  theme_void() +
  ggtitle("Merge all samples") + 
  theme(plot.title = element_text(hjust = 0.5)) +  NoLegend()
``` 

# Feature plot
``` {r, fig.width= 7, fig.height= 5}

FeaturePlot(All_arthritis_integration_TSNE, features = c("CD3E", "CD8A", "CD4", "FOXP3"), pt.size = 0.2,
    ncol = 2, reduction = "tsne")
```


``` {r, fig.width= 8, fig.height= 11}
FeaturePlot(All_arthritis_integration_TSNE, features = c("EPCAM", "PTPRC", "COL1A1", "CD79A", "CD3D", "S100A8", "S100A9", "CD8A"), pt.size = 0.2,
    ncol = 2)
```

``` {r}
sessionInfo()
```

