---
title: "Arthritis single cell-- Clustering"
date: "2023-11-08"
output:
  html_document:
    toc: yes
    toc_depth: 2
    number_sections: yes
    toc_float:
      collapsed: yes
---

# Load library 
``` {r, results='hide'}
library(tidyverse)
library(Seurat)
library(harmony)
set.seed(12345)
```

# Objective

This section will focus on how to fine-tune UMAP plot. 


# Integration using harmony

``` {r, results='hide'}
#read the PCA object back
Arthritis<- readRDS("Arthritis_PCA.RDS")
Arthritis<- RunHarmony(Arthritis ,group.by.vars = "orig.ident")
```

# Find Neighbors and clusters

``` {r, fig.width= 8, fig.height= 10, results='hide'}
set.seed(12345)
Arthritis.clustree<- FindNeighbors(Arthritis, reduction = "harmony", dims = 1:30, verbose = FALSE)

#run resolution between 0 and 1
resolution_values <- seq(0, 1, by = 0.1)
Arthritis.clustree<- FindClusters(Arthritis.clustree, resolution = resolution_values, random.seed = 12345, verbose = FALSE)
``` 

# Examine cluster tree to deterine the optimal resolution 

``` {r, fig.width= 8, fig.height= 9}
library(clustree)
clustree(Arthritis.clustree)
```

# Investigate various min.dist value of RunUMAP function

``` {r, results='hide'}
# Determine the optimal min.dist values 
min_dist_values <- c(0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.1, 0.05, 0.01, 0.005, 0.001)  

# Create a list to store the plots
plot_list <- list()

#Iterate through all these min.dist values
for(dist in min_dist_values){
    Arthritis.min.dist <- RunUMAP(object = Arthritis ,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = dist)

    Arthritis.min.dist <- FindNeighbors(object = Arthritis.min.dist, reduction="harmony", dims = 1:30)
    snn.obj.min.dist <- FindClusters(object = Arthritis.min.dist, resolution = 0.2, random.seed = 12345)
    
    umap_plot.min.dist <- DimPlot(snn.obj.min.dist, reduction="umap", ncol=2, group.by = "seurat_clusters", label = TRUE, label.size = 3) +
      ggtitle(paste("(min.dist =", dist, ")")) + NoLegend()
    
    # Append the plot to the list
    plot_list[[as.character(dist)]] <- umap_plot.min.dist
}
```


# Combine all plots into a single figure.
``` {r,fig.width= 12, fig.height=15}
library(ggplot2)
library(gridExtra)

# Calculate the number of rows and columns for the grid
total_plots.dist <- length(min_dist_values)
n_rows.dist <- ceiling(sqrt(total_plots.dist))
n_cols.dist <- ceiling(total_plots.dist / n_rows.dist)

# Combine all plots into a flexible grid
combined_plot.dist <- do.call(grid.arrange, c(plot_list, ncol = n_cols.dist, nrow = n_rows.dist))
```

## Save the combined plot
``` {r}
ggsave("combined_min_dist_plots.png", combined_plot.dist, height=18, width=20)
``` 

Comments: I have saved all these min.dist plots and can be viewed outside Rmarkdown. 


# Examine various resolution values

``` {r, results='hide'}
# Create a list to store the plots
plot_list_resolution <- list()

# Run UMAP
Arthritis.res<- RunUMAP(object = Arthritis,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = 0.05)
Arthritis.res <- FindNeighbors(object =  Arthritis.res, reduction="harmony", dims = 1:30)

#iterate through various value of resolutions 
for(res in c(0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)){
    snn.obj.res <- FindClusters(object = Arthritis.res, resolution = res, random.seed = 12345, verbose = FALSE)
    
    umap_plot_res <- DimPlot(snn.obj.res, reduction="umap", ncol=2, group.by = "seurat_clusters", label = TRUE, label.size = 3) +
      NoLegend() +
      ggtitle(paste("(resolution =", res,")"))
    
    # Append the plot to the list
    plot_list_resolution[[as.character(res)]] <- umap_plot_res 
}
``` 

# Combine all plots into a single figure.
``` {r, fig.width= 9, fig.height=11}
library(ggplot2)
library(gridExtra)

# Calculate the number of rows and columns for the grid
total_plots <- length(c(0.01, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1))
n_rows <- ceiling(sqrt(total_plots))
n_cols <- ceiling(total_plots / n_rows)

# Combine all plots into a flexible grid
combined_plot_res <- do.call(grid.arrange, c(plot_list_resolution, ncol = n_cols, nrow = n_rows))
``` 

# Save the combined plot
``` {r}
ggsave("combined_resolution_plots.png", combined_plot_res, height=18, width=20,  dpi=300)
``` 


# Run single umap with optimal resolution and min.dist

Here I test resolution= 0.5 and min.dist= 0.01

``` {r, fig.width= 5, fig.height=3}
umap<- RunUMAP(object = Arthritis,
                          reduction = "harmony",
                          dims = 1:30,
                          min.dist = 0.01)
umap <- FindNeighbors(object =  umap, reduction="harmony", dims = 1:30)
umap<- FindClusters(object = umap , resolution = 0.5 , random.seed = 12345, verbose = FALSE)
DimPlot(umap, reduction="umap", label = TRUE, label.size = 3) +
      theme_minimal() +
  NoLegend()

```

# Locate Treg cluster

``` {r, fig.width= 7, fig.height= 6}
# These are now standard steps in the Seurat workflow for visualization and clustering
# Visualize canonical marker genes as violin plots.
VlnPlot(umap, features = c("CD3E", "CD8A", "CD4", "FOXP3"),
    pt.size = 0.2, ncol = 2)

``` 
# Feature plots

``` {r, fig.width= 7, fig.height= 5}
# Visualize canonical marker genes on the sctransform embedding.
FeaturePlot(umap, features = c("CD3E", "CD8A", "CD4", "FOXP3"), pt.size = 0.2,
    ncol = 2)
``` 

# Look like with resolution= 0.5 and min.dist= 0.01, Treg is well separated from other cell cluster. 

``` {r}
sessionInfo()
```

